\chapter{Pebbling Comonad}
\section{Introduction}
The Pebbling game, similar in structure to the Ehrenfreucht-{\Fraisse} game, is used to prove equivalence between structures in the $k$-variable fragments of infinitary logic. Given two structures $A,B$, the $k$-pebble game, denoted $\pgameSym{k}{A}{B}$, is played with Spoiler and Duplicator both having a set of $k$-pebbles $[k] = \{1,\dots,k\}$.  
\begin{itemize} 
\item Spoiler places one of his pebbles $p \in [k]$ on an element in either structure $a \in A$ or $b \in B$. If pebble $p$ is already placed on a different element, then Spoiler removes the pebble from this element and places the pebble on the new element he chose.  
\item Duplicator places her corresponding pebble $p \in [k]$ on an element in the other structure $b \in B$ or $a \in A$. Just as with Spoiler, Duplicator may have to move her pebble from a previously pebbled element.
\end{itemize} 
The game is played for $\omega$ many rounds. At the end of the game, $(a_{1},\dots,a_{k})$ and $(b_{1},\dots,b_{k})$ are the pebbled elements. Duplicator wins the $k$-pebble game if the map $\gamma:a_{p} \longmapsto b_{p}$ is a partial $\sigma$-isomorphism from $A$ to $B$. Otherwise, Spoiler wins. Just as with the case with EF games, the asymmetric (or existential positive) game from $A$ to $B$, denoted $\pgame{k}{A}{B}$, is the same game with the additional restriction that Spoiler must always pebble elements in $A$ and that the map obtained is a partial $\sigma$-morphism. The following result holds:
\begin{prop}
The following are equivalent:
\begin{itemize}
\item Duplicator has a winning strategy in $\pgameSym{k}{A}{B}$
\item $\equivL{A}{B}{\mathcal{L}^{k}_{\infty,\omega}}$, i.e. for every $\phi \in \mathcal{L}^{k}_{\infty,\omega}$, $A \vDash \phi \Leftrightarrow B \vDash \phi$
\end{itemize}
\end{prop}
As was the goal with the EF game comonad, our goal is to construct a $\sigma$-structure $\pcomonad{k}{A}$ from a $\sigma$-structure $A$, that ``internalizes'' the asymmetric and symmetric $k$-pebble games in the category $\mathcal{R}(\sigma)$. The construction of $\pcomonad{k}{}$ and all of the results in this chapter were originally discovered in \cite{Abramsky2017}. The paper \cite{Abramsky2017} was the inspiration to develop the other game comonads in this dissertation. We restate the propositions here to demonstrate the connections and differences with these other comonads.  
\section{Comonad laws}
Let $A$ be a $\sigma$-structure over relational signature $\sigma$, then for every $k \in \mathbb{N}$ we define a $\sigma$-structure $\pcomonad{k}{A}$. Intuitively, $\pcomonad{k}{A}$ is the structure of the finite $k$-pebblings on $A$. Let the domain of this structure be $|\pcomonad{k}{A}| = ([k] \times A)^{<\omega}$ (i.e. the set of finite sequence of elements in product $[k] \times A$). In order, to define the $\sigma$-structure on $\pcomonad{k}{A}$, a bit more notation is necessary:
\begin{defn}
Define, for every $\sigma$-structure $A$, $\epsilon_{A}:\pcomonad{k}{A} \longrightarrow A$ by $[(p_{1},a_{1}),\dots,(p_{n},a_{n})] \mapsto a_{n}$ (i.e. the element of the last move of the play). Further, define $\pi_{A}:\pcomonad{k}{A} \longrightarrow \{1,\dots,k\}$ by $[(p_{1},a_{1}),\dots,(p_{n},a_{n})] \mapsto p_{n}$ (i.e. the pebble index of the last move of the play).
\label{defn:epsilonP}
\end{defn}
With this definition in place, we can define the $\sigma$-structure on $\pcomonad{k}{A}$. Suppose $R \in \sigma$ is an $m$-ary relation, then we define the interpretation $R^{\pcomonad{k}{A}}$ such that for every $s_{1},\dots,s_{m} \in |\pcomonad{k}{A}|$, 
\begin{align}
(s_{1},\dots,s_{m}) \in R^{\pcomonad{k}{A}}  &\Leftrightarrow    \text{ for every $i,j$, } s_{i} \sqsubseteq s_{j} \text{ or } s_{j} \sqsubseteq s_{i} \text{(i.e. exists $\sqsubseteq$-greatest $s_{*}$)} & \label{eq:R1st}\\
&\text{ and for every $i$, } \pi_{A}(s_{i}) \text{ does not appear in suffix of $s_{i}$ in $s_{*}$ } & \label{eq:R2nd} \\
&\text{ and } (\epsilon_{A}(s_{1}),\dots,\epsilon_{A}(s_{m})) \in R^{A} & \label{eq:R3rd}
\end{align}
The definition of $\pcomonad{k}{A}$ can be extended to morphisms of $\sigma$-structures. 
\begin{defn}
Given a morphism $f:A \longrightarrow B$, define the morphism $\pcomonad{k}{f}:\pcomonad{k}{A} \longrightarrow \pcomonad{k}{B}$ by $[(p_{1},a_{1}),\dots,(p_{n},a_{n})] \mapsto [(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))]$
\label{defn:comonadMorphismP}
\end{defn}
\begin{prop}
The definition (\ref{defn:comonadMorphismP}) of $\pcomonad{k}{f}:\pcomonad{k}{A} \longrightarrow \pcomonad{k}{B}$ given above is indeed a morphism of $\sigma$-structures. 
\begin{proof}
Suppose $R \in \sigma$, then we want to show that if $(s_{1},\dots,s_{m}) \in R^{\pcomonad{k}{A}}$, then \linebreak $(\pcomonad{k}{f}(s_{1}),\dots,\pcomonad{k}{f}(s_{m})) \in R^{\pcomonad{k}{B}}$. For brevity, assume that $R$ is a binary relation (the proof for a general $m$-ary relation is a straightforward generalization). Suppose $s,s' \in \pcomonad{k}{A}$ such that $(s,s') \in R^{\pcomonad{k}{A}}$. Let $s = [(p_{1},a_{1}),\dots,(p_{n},a_{n})]$ and $s' = [(q_{1},a_{1}),\dots,(q_{m},a'_{m})]$. We aim to show that $(\pcomonad{k}{f}(s),\pcomonad{k}{f}(s')) \in R^{\pcomonad{k}{B}}$ \\
\begin{enumerate}
\item  Since $(s,s') \in R^{\pcomonad{k}{A}}$, by condition (\ref{eq:R1st}), $s \sqsubseteq s'$ or $s' \sqsubseteq s$. Without loss of generality, assume $s \sqsubseteq s'$. Since $s \sqsubseteq s'$.
$$s' = [(p_{1},a_{1}),\dots,(p_{n},a_{n}),(q_{n+1},a'_{n+1}),\dots,(q_{m},a'_{m})]$$ 
(noting that for $i \leq n$, $p_{i} = q_{i}$ and $a_{i} = a'_{i}$). Therefore $$\pcomonad{k}{f}(s) = [(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))]$$ 
and 
$$\pcomonad{k}{f}(s') =[(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n})),(q_{n+1},f(a'_{n+1})), \dots,(q_{m},f(a'_{m}))]$$ 
Hence, $\pcomonad{k}{f}(s) \sqsubseteq \pcomonad{k}{f}(s')$ and (\ref{eq:R1st}) is satisfied. 
\item By condition (\ref{eq:R2nd}) and $(s,s') \in R^{\pcomonad{k}{f}}$, for $n < i \leq m$, $p_{n} \not= q_{i}$. Since $\pcomonad{k}{f}$ does not affect pebble indices, (\ref{eq:R2nd}) is satisfied.
\item  By condition (\ref{eq:R3rd}) and $(s,s') \in R^{\pcomonad{k}{f}}$, $(\epsilon_{A}(s),\epsilon_{A}(s')) = (a_{n},a'_{m}) \in R^{A}$. Since $f:A \rightarrow B$ is a morphism of $\sigma$-structures, $(f(a_{n}),f(a'_{m})) \in R^{B}$. That is, $(\epsilon_{B}\circ \pcomonad{k}{f}(s),\epsilon_{B} \circ \pcomonad{k}{f}(s')) \in R^{B}$. Hence, (\ref{eq:R3rd}) is satisfied.
\end{enumerate}
Therefore, $(\pcomonad{k}{f}(s),\pcomonad{k}{f}(s')) \in R^{\pcomonad{k}{B}}$ and $\pcomonad{k}{f}$ is indeed a morphism of $\sigma$-structures. 
\end{proof}
\end{prop}
\begin{prop}
$\epsilon:\pcomonad{k}{} \longrightarrow 1_{\mathcal{R}(\sigma)}$ is a natural transformation.
\begin{proof}
For every $A,B \in \mathcal{R}(\sigma)$ we want to show that:
\begin{equation}
\bfig \square[\pcomonad{k}{A}`A`\pcomonad{k}{B}`B;\epsilon_{A}`\pcomonad{k}{f}`f`\epsilon_{B}] \efig
\label{eq:epsilonNP}
\end{equation}
\begin{align*}
f \circ \epsilon_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})])   &= f(a_{n}) & \text{defn (\ref{defn:epsilonP}) $\epsilon_{A}$}\\
&= \epsilon_{B}([(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))]) & \text{defn (\ref{defn:epsilonP}) $\epsilon_{B}$}\\
&= \epsilon_{B} \circ \pcomonad{k}{f}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{f}$}
\end{align*}
Hence, the above diagram (\ref{eq:epsilonNP}) commutes.
\end{proof}
\label{prop:epsilonNP}
\end{prop}
\begin{defn}
Suppose $s \in \pcomonad{k}{A}$, then $s = [(p_{1},a_{1}),\dots,(p_{n},a_{n})]$ for some $n \in \omega$ and for every $i = 1,\dots, n$, $p_{i} \in \{1,\dots,k\}$, $a_{i} \in A$. Let $s_{i} = [(p_{1},a_{1}),\dots,(p_{i},a_{i})] \in \pcomonad{k}{A}$. Define, for every $\sigma$-structure $A$, $\delta_{A}:\pcomonad{k}{A} \longrightarrow \pcomonad{k}{\pcomonad{k}{A}}$ by $s \mapsto [(p_{1},s_{1}),\dots,(p_{n},s_{n})]$.
\label{defn:deltaP}
\end{defn}
\begin{prop}
$\delta:\pcomonad{k} \longrightarrow \pcomonad{k}{\pcomonad{k}{}}$ is a natural transformation.
\begin{proof}
For every $A,B \in \mathcal{R}(\sigma)$ we want to show that:
\begin{equation}
\bfig \square[\pcomonad{k}{A}`\pcomonad{k}{\pcomonad{k}{A}}`\pcomonad{k}{B}`\pcomonad{k}{\pcomonad{k}{B}};\delta_{A}`\pcomonad{k}{f}`\pcomonad{k}{\pcomonad{k}{f}}`\delta_{B}] \efig
\label{eq:deltaNP}
\end{equation}
\begin{alignat*}{2}
\pcomonad{k}{\pcomonad{k}{f}} \circ \delta_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})])   &= \pcomonad{k}{\pcomonad{k}{f}}([(p_{1},s_{1}),\dots,(p_{n},s_{n})]) & \text{defn (\ref{defn:deltaP}) $\delta_{A}$} \\
&= [(p_{1},\pcomonad{k}{f}(s_{1})),\dots,(p_{n},\pcomonad{k}{f}(s_{n}))] & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{\pcomonad{k}{f}}$}\\
&= [(p_{1},[(p_{1},f(a_{1}))]),\dots,(p_{n},[(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))])] & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{f}$}\\
&= \delta_{B}([(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))]) & \text{defn (\ref{defn:deltaP}) $\delta_{B}$} \\
&= \delta_{B} \circ \pcomonad{k}{f}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{f}$}
\end{alignat*}
Hence, the above diagram (\ref{eq:deltaNP}) commutes.
\end{proof}
\label{prop:deltaNP}
\end{prop}
\begin{thm}
The triple $\langle \pcomonad{k}{},\delta,\epsilon \rangle$ is a comonad.
\begin{proof}
By proposition (\ref{prop:deltaNP}) and (\ref{prop:epsilonNP}), $\delta$ and $\epsilon$ are natural transformations. Hence, $\delta$ and $\epsilon$ are indeed the comultiplication and counit of $\pcomonad{k}{}$. The associative and identity laws remain to be shown. \\
For associativity, for every $A \in \mathcal{R}(\sigma)$, the following diagram commutes:  
\begin{equation}
\bfig \Square[\pcomonad{k}{A}`\pcomonad{k}{\pcomonad{k}{A}}`\pcomonad{k}{\pcomonad{k}{A}}`\pcomonad{k}{\pcomonad{k}{\pcomonad{k}{A}}};\delta_{A}`\delta_{A}`\delta_{\pcomonad{k}{A}}`\pcomonad{k}{\delta_{A}}] \efig 
\end{equation}
\begin{center}
\begin{alignat*}{2}
\delta_{\pcomonad{k}{A}} \circ \delta_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})])   &= \delta_{\pcomonad{k}{A}}([(p_{1},s_{1}),\dots,(p_{n},s_{n})]) & \text{defn (\ref{defn:deltaP}) $\delta_{A}$} \\
&= [(p_{1},[(p_{1},s_{1})]),\dots,(p_{n},[(p_{1},s_{1}),\dots,(p_{n},s_{n})])]  & \text{defn (\ref{defn:deltaP}) $\delta_{\pcomonad{k}{A}}$}\\
&= [(p_{1},\delta_{A}(s_{1})),\dots,(p_{n},\delta_{A}(s_{n}))] & \text{defn (\ref{defn:deltaP}) $\delta_{A}$}  \\
&= \pcomonad{k}{\delta_{A}}([(p_{1},s_{1}),\dots,(p_{n},s_{n})]) & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{\delta_{A}}$}  \\
&= \pcomonad{k}{\delta_{A}} \circ \delta_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) & \text{defn (\ref{defn:deltaP}) $\delta_{A}$}\\
\end{alignat*}
\end{center}
For identity, for every $A \in \mathcal{R}(\sigma)$, the following diagram commutes:  
\begin{equation}
\bfig 
    \square[\pcomonad{k}{A}`\pcomonad{k}{\pcomonad{k}{A}}`\pcomonad{k}{\pcomonad{k}{A}}`\pcomonad{k}{A};\delta_{A}`\delta_{A}`\pcomonad{k}{\epsilon_{A}}`\epsilon_{\pcomonad{k}{A}}] 
    \morphism(0,500)/=/<500,-500>[\pcomonad{k}{A}`\pcomonad{k}{A};]
\efig 
\end{equation}
\begin{align*}
\pcomonad{k}{\epsilon_{A}} \circ \delta_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) &= \pcomonad{k}{\epsilon_{A}}([(p_{1},s_{1}),\dots,(p_{n},s_{n})]) & \text{defn (\ref{defn:deltaP}) $\delta_{A}$}\\
&= [(p_{1},\epsilon_{A}(s_{1})),\dots,(p_{n},\epsilon_{A}(s_{n}))]  & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{\epsilon_{A}}$}  \\
&= [(p_{1},a_{1}),\dots,(p_{n},a_{n})] & \text{defn (\ref{defn:epsilonP}) $\epsilon_{A}$}\\
\hspace{1pt}\\
&= s_{n} & \text{defn (\ref{defn:deltaP}) $s_{n}$}\\
&= \epsilon_{\pcomonad{k}{A}}([(p_{1},s_{1}),\dots,(p_{n},s_{n})]) & \text{defn (\ref{defn:epsilonP}) $\epsilon_{\pcomonad{k}{A}}$} \\
&= \epsilon_{\pcomonad{k}{A}} \circ \delta_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) & \text{defn (\ref{defn:deltaP}) $\delta_{A}$}
\end{align*}
By definition, $\pcomonad{k}{}$ is a comonad.
\end{proof}
\end{thm}
For every $l,k \in \omega$ such that $l \leq k$ and $\sigma$-structure $A$, there exists an inclusion $i_{A}^{l,k}: \pcomonad{l}{A} \longrightarrow \pcomonad{k}{A}$. 
\begin{prop}
The inclusion maps form a natural transformation $i^{l,k}:\pcomonad{l}{} \longrightarrow \pcomonad{k}{}$. Further, each map preserves the counit and comultiplication (i.e. each map is a morphism of comonads). 
\end{prop}
\begin{proof}
\begin{equation}
\bfig \square[\pcomonad{l}{A}`\pcomonad{k}{A}`\pcomonad{l}{B}`\pcomonad{k}{B};i^{l,k}_{A}`\pcomonad{l}{f}`\pcomonad{k}{f}`i^{l,k}_{B}]\efig
\end{equation}
\begin{align*}
\pcomonad{k}{f} \circ i^{l,k}_{A}([(p_{1},a_{1}),\dots,(p_{n},a_{n})])   &= \pcomonad{k}{f}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) & p_{i} \in \{1,\dots, l\} \subseteq \{1,\dots,k\}\\
&= [(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))] &\text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{f}$} \\
&= i^{l,k}_{B}([(p_{1},f(a_{1})),\dots,(p_{n},f(a_{n}))]) & p_{i} \in \{1,\dots, l\} \subseteq \{1,\dots,k\}\\
&= i^{l,k}_{B} \circ \pcomonad{l}{f}([(p_{1},a_{1}),\dots,(p_{n},a_{n})]) & \text{defn (\ref{defn:comonadMorphismP}) $\pcomonad{k}{f}$}\\
\end{align*}
\end{proof}
The grading given by these inclusion maps seem to suggest that there is a colimit object capturing the information of $\pcomonad{k}{A}$ for every $k \in \omega$. This is indeed the case. Consider the structure $\pcomonad{\omega}{A}$ with domain $|\pcomonad{\omega}{A}| = (\omega \times A)^{\omega}$. The structure on $\pcomonad{\omega}{A}$ is similar to the structure given to $\pcomonad{k}{A}$. 
\begin{prop}
Let $\omega$ be considered as a poset category under the usual order. The object $\pcomonad{\omega}{A}$ is the $\omega$-colimit of the family $\{\pcomonad{k}{A}\}_{k \in \omega}$ with the above inclusion maps. 
\begin{proof}
For every $k \in \omega$, define $i^{k}_{A}:\pcomonad{k}{A} \rightarrow \pcomonad{\omega}{A}$ as the inclusion (i.e. $[(p_{1},a_{1}),\dots,(p_{n},a_{n})] \mapsto [(p_{1},a_{1}),\dots,(p_{n},a_{n})]$). Clearly, the following diagram commutes for all $l,k \in \omega$ with $l \leq k$
\begin{equation}
\bfig \Vtriangle[\pcomonad{l}{A}`\pcomonad{k}{A}`\pcomonad{\omega}{A};i^{l,k}_{A}`i^{l}_{A}`i^{k}_{A}]\efig
\label{eq:omegaColimitP}
\end{equation}
Suppose that there exists a $\sigma$-structure $B$ and for every $l,k \in \omega$ with $l \leq k$, there exist morphisms $f^{l}:\pcomonad{l}{A} \longrightarrow B$, $f^{k}:\pcomonad{k}{A} \longrightarrow B$ such that $f^{l} = f^{k} \circ i^{l,k}$. Consider the morphism $u:\pcomonad{\omega}{A} \longrightarrow B$ given by $[(p_{1},a_{1}),\dots,(p_{n},a_{n})] \mapsto f^{j}([(p_{1},a_{1}),\dots,(p_{n},a_{n})])$ where $j = \max(p_{1},\dots,p_{n})$ 
\begin{equation}
\bfig 
    \Vtriangle[\pcomonad{l}{A}`\pcomonad{k}{A}`\pcomonad{\omega}{A};i^{l,k}_{A}`i^{l}_{A}`i^{k}_{A}]
    \morphism(0,500)|l|/{@{>}@/^-7pt/}/<500,-1000>[\pcomonad{l}{A}`B;f^{l}]
    \morphism(1000,500)|r|/{@{>}@/^7pt/}/<-500,-1000>[\pcomonad{k}{A}`B;f^{k}]
    \morphism(500,0)|m|/.>/<0,-500>[\pcomonad{\omega}{A}`B;u]
\efig
\label{eq:omegaColimitUP}
\end{equation}
Moreover, given the conditions on $f^{j}$ for all $j \in \omega$, $u$ is unique. Namely, suppose there exists a morphism $u':\pcomonad{\omega}{A} \longrightarrow B$ such that for all $j \in \omega$, $f^{j} = u' \circ i^{j}_{A}$. Suppose $s = [(p_{1},a_{1}),\dots,(p_{n},a_{n})] \in \pcomonad{\omega}{A}$ and $k = \max(p_{1},\dots,p_{n})$, then for all $j \geq k, s \in \pcomonad{j}{A}$.  
\begin{align*}
u(s)    &= f^{k}(s) & \text{defn of $u$} \\
        &= f^{j} \circ i^{k,j}_{A}(s) & \text{fig. (\ref{eq:omegaColimitUP})} \\
        &= u' \circ i^{j}_{A} \circ i^{k,j}_{A}(s) & \text{hypothesis on $u'$} \\
        &= u' \circ i^{k}_{A}(s) & \text{fig. (\ref{eq:omegaColimitP})}\\
        &= u'(s) & \text{inclusion} 
\end{align*}
Since $u(s) = u'(s)$ for all $s \in \pcomonad{\omega}{A}$, $u = u'$ so $u$ is unique as desired.  
\end{proof}
\end{prop}
\section{Positional Form and Equivalences}\label{sec:positionalFormP}
In order to connect the construction $\pcomonad{k}{}$ and the pebbling game, we have to encode strategies in the game into positional form, just as in the case for the EF game. \\~\\
Let $\Gamma_{k}(A,B)$ be the set of all partial functions $\gamma:[k] \rightharpoondown A \times B$. Intuitively, $\gamma(q) = (a,b)$ represents that position that the pebble $q \in [k]$ is on elements $a \in A$ and $b \in B$. We will need a different update operation than one that was used for sequences in the positional form for the EF game. Let $\gamma' = \gamma[p \mapsto (a,b)]$ be the partial function such that:  
\begin{equation*}
\gamma'(q) = \begin{cases}
                \gamma(q) & \text{for } q \not= p \\
                (a,b) & \text{for } q = p 
            \end{cases}
\end{equation*}
In this case, the update operation is denoted $\gamma \xrightarrow{(p,a):b} \gamma'$. We will use $\gamma \longrightarrow \gamma'$ if $\gamma \xrightarrow{(p,a):b} \gamma'$ for some $p \in [k], a \in A$ and $b \in B$.
\begin{defn}
$S \subseteq \Gamma_{k}(A,B)$ is a \textit{strategy in positional form} if $S$ satisfies the following conditions:
\begin{enumerate}[label=(S\arabic*),ref=S\arabic*,start=0]
\item $\emptySeq \in S$ \label{eq:S1st}
\item For all $\gamma \in S$, $a \in A$ and $p \in [k]$ there exists a $b \in B$ such that $\gamma[p \mapsto (a,b)] \in S$ \label{eq:S2nd}
\item $S$ is reachable: For all $\gamma \in S$, there is a chain \label{eq:S3rd}
$$\gamma_{0} \longrightarrow \dots \longrightarrow \gamma_{n}$$
such that $\gamma_{0} = \emptySeq$, $\gamma_{n} = \gamma$ and $\gamma_{i} \in S$. 
\end{enumerate}
\end{defn}
Given a function $f:\pcomonad{k}{A} \longrightarrow B$, define $\theta_{f}:\pcomonad{k}{A} \longrightarrow \Gamma_{k}(A,B)$ by $s = [(p_{1},a_{1},\dots,(p_{n},a_{n})] \mapsto \gamma$ with $\gamma$ the partial function from $p_{i_{1}},\dots,p_{i_{n}}$ (the distinct pebbles among $p_{1},\dots,p_{n}$ to the pairs $(a_{i_{j}},b_{i_{j}})$ where $a_{i_{j}}$ is the last element in $s$ pebbled by $p_{i_{j}}$ and $b_{i_{j}}$ is the corresponding element in $f^{*}(s)$). 
\begin{prop}[{\cite[Proposition 9]{Abramsky2017}}]
If $f:\pcomonad{k}{A} \longrightarrow B$ is a $\sigma$-morphism, then there exists a strategy in positional form $S_{f}$.
\begin{proof}
Just as with the EF comonad, let $S_{f}$ be the image of $\theta_{f}$ with $\emptySeq$ added. It is easy to check that the properties (\ref{eq:S1st})-(\ref{eq:S3rd}) are satisfied.
\end{proof}
\label{prop:fToPosFormP}
\end{prop}
\begin{prop}[{\cite[Proposition 9]{Abramsky2017}}]
Conversely, every strategy in positional form $S$ there exists a $f:\pcomonad{k}{A} \rightarrow B$ such that $S = S_{f}$
\begin{proof}
This proof is more complicated than (\ref{prop:posFormToFEF}). Namely, $\pcomonad{k}{A}$ is a infinite set and $\Gamma_{k}(A,B)$ is finite. Multiple plays can lead to the same position. To resolve this non-determinism, we fix a linear order $<$ of $B$ and ensure that $f$ is constructed to respond to the $<$-least $b$ given by the strategy $S$ that does not already appear in the play. The full details are in \cite[Proposition 9]{Abramsky2017}.
\end{proof}
\label{prop:posFormToFP}
\end{prop}
\subsection{Equivalence $\exists^{+}\mathcal{L}^{k}_{\infty,\omega}$}
Let $R(\gamma)$ denote the image of $\gamma$ in $A \times B$.
\begin{defn}
A position $\gamma$ is \textit{winning for Duplicator in $\pgame{k}{A}{B}$} if $R(\gamma)$ is a partial $\sigma$-morphism from $A$ to $B$.  
Naturally, we can extend the definition to say that a strategy in positional form $S \subseteq \Gamma_{k}(A,B)$ is \textit{winning for Duplicator in $\pgame{k}{A}{B}$} if for all $\gamma \in S$, $\gamma$ is winning for Duplicator. Function $f:\pcomonad{k}{A} \longrightarrow B$ is \textit{winning for Duplicator in $\pgame{k}{A}{B}$} if $S_{f}$ is winning for Duplicator in $\pgame{k}{A}{B}$.
\end{defn}
Just as with the EF game, we have to lift $\sigma$-structures to expanded signature $\sigma' = \sigma \cup \{I\}$ with $I$ a binary relation. Recall that $\sigma'$-structure $A$ is pure if $I$ is just the identity relation. Note that if $A$ is a pure $\sigma'$-structure, then by the definition of $\pcomonad{k}{A}$ as a $\sigma'$-structure, $\pcomonad{k}{A}$ is not a pure $\sigma'$-structure. However, two plays are $I$-related if the same pebble is not on distinct elements. This ensures that $S_{f}$ contains well-defined partial functions.
\begin{thm}[{\cite[Proposition 10]{Abramsky2017}}]
If $A,B$ are pure $\sigma'$-structures and $f:\pcomonad{k}{A} \rightarrow B$ is a function, then
\center{$f:\pcomonad{k}{A} \longrightarrow B$ is a $\sigma'$-morphism if and only if $f$ is winning for Duplicator in $\pgame{k}{A}{B}$}
\begin{proof}
The proof is essentially the same as the one given for (\ref{thm:toPositionalFormEF}). 
\end{proof}
\begin{cor}[]
There exists a $\sigma'$-morphism $f:\pcomonad{k}{A} \longrightarrow B$ if and only if for all $\phi \in \exists^{+}\mathcal{L}^{k}_{\infty,\omega}, A \vDash \phi \Rightarrow B \vDash \phi$
\label{cor:forthOneP}
\end{cor}
\begin{cor}
There exists $\sigma'$-morphisms $f:\pcomonad{k}{A} \longrightarrow B$, $g:\pcomonad{k}{B} \longrightarrow A$ if and only if $\equivL{A}{B}{\exists^{+}\mathcal{L}^{k}_{\infty,\omega}}$. 
\begin{proof}
Recall definition (\ref{defn:equivLogic}) and apply above corollary (\ref{cor:forthOneP}) to $f$ and $g$.  
\end{proof}
\label{cor:forthP}
\end{cor}
\label{thm:toPositionalFormP}
\end{thm}
\subsection{Equivalence $\mathcal{L}^{k}_{\infty,\omega}$}
The comonad $\pcomonad{k}{}$ also captures equivalence $\equivL{A}{B}{\mathcal{L}^{k}_{\infty,\omega}}$ in the full $k$-variable fragment of infinitary logic. Just as with the EF game, to formalize a winning strategy for Duplicator in the symmetric $k$-pebbling game $\pgameSym{k}{A}{B}$ we need winning strategies for Duplicator in $\pgame{k}{A}{B}$ given by $\sigma'$-morphism $f:\pcomonad{k}{A} \longrightarrow B$, and $\pgame{k}{B}{A}$ given by $\sigma'$-morphism $\pcomonad{k}{B} \longrightarrow A$ that are compatible in positional form. Just as with the EF game, we need to define a `swap' monotone function $()^{-}:\Gamma_{k}(A,B) \longrightarrow \Gamma_{k}(B,A)$. Given $\gamma \in \Gamma_{k}(A,B)$, define $\gamma^{-} \in \Gamma_{k}(B,A)$ such that $\gamma^{-}(p) = (b,a)$ iff $\gamma(p) = (a,b)$. 
\begin{defn}
A strategy in positional form $S \subseteq \Gamma_{k}(A,B)$ is \textit{winning for Duplicator in $\pgameSym{k}{A}{B}$} if $S$ is winning for Duplicator in $\pgame{k}{A}{B}$ and $S^{-}$ is winning for Duplicator in $\pgame{k}{B}{A}$. Naturally, we can extend this definition to functions, $f:\pcomonad{k}{A} \longrightarrow B$ is \textit{winning for Duplicator in $\pgameSym{k}{A}{B}$} if $S_{f}$ is winning for Duplicator in $\pgameSym{k}{A}{B}$.
\end{defn}
It follows by (\ref{prop:posFormToFP}), that if $f$ is winning for Duplicator in $\pgameSym{k}{A}{B}$, there exists a function $g:\pcomonad{k}{B} \longrightarrow A$ such that $S_{g}$ is winning for Duplicator in $\pgame{k}{B}{A}$ and $S_{f}^{-} = S_{g}$, so $g$ is winning for Duplicator in $\pgameSym{k}{A}{B}$. This is the right definition to characterize equivalence in $\mathcal{L}^{k}_{\infty,\omega}$. 
\begin{prop}[{\cite[Theorem 20]{Abramsky2017}}]
For all $k \in \omega$, the following are equivalent:
\begin{enumerate}[label=(\arabic*)$_{k}$]
\item $\equivL{A}{B}{\mathcal{L}^{k}_{\infty,\omega}}$, i.e. for every $\phi \in \mathcal{L}^{k}_{\infty,\omega}$, $A \vDash \phi \Leftrightarrow B \vDash \phi$
\item There exists a strategy in positional form $S \subseteq \Gamma_{k}(A,B)$ such that $S$ is winning for Duplicator in $\pgameSym{k}{A}{B}$
\end{enumerate}
\begin{proof}
Since for all $\gamma \in S$ is winning for Duplicator in $\pgame{k}{A}{B}$ and $\gamma^{-} \in S^{-}$ is winning for Duplicator in $\pgame{k}{B}{A}$ means that $R(\gamma)$ is a partial isomorphism. By results in \cite{Immerman1982}, the proposition holds.
\end{proof}
\begin{cor}
\begin{enumerate}[label=(\arabic*)]
\item $\equivL{A}{B}{\mathcal{L}^{k}_{\infty,\omega}}$ 
\item There exist $\sigma'$-morphisms $f:\pcomonad{k}{A} \longrightarrow B$, $g:\pcomonad{k}{B} \longrightarrow A$ such that $S_{f} = S_{g}^{-}$ 
\item There exists $f:\pcomonad{k}{A} \longrightarrow B$ winning for Duplicator in $\pgameSym{k}{A}{B}$ 
\end{enumerate}
\begin{proof}
$(1) \Rightarrow (2)$ By proposition (\ref{prop:backForthSyntaxP}), there exists a strategy in positional form $S$ winning for Duplicator in $\pgameSym{k}{A}{B}$. Hence, $S$ winning for Duplicator in $\pgame{k}{A}{B}$ and $S^{-}$ winning for Duplicator in $\pgame{k}{B}{A}$. By theorem (\ref{thm:toPositionalFormP}), there exists $\sigma'$-morphisms $f:\pcomonad{k}{A} \longrightarrow B$, $g:\pcomonad{k}{B} \longrightarrow A$ with $S_{f} = S$, and $S^{-} = S_{g}$. \\ 
$(2) \Rightarrow (3)$ Straightforward from definition of $f$ winning for Duplicator in $\pgameSym{k}{A}{B}$. \\ 
$(3) \Rightarrow (1)$ By proposition (\ref{prop:backForthSyntaxP}) with $S_{f}$ witnessing $S$.  
\end{proof}
\label{cor:backForthP}
\end{cor}
\label{prop:backForthSyntaxP}
\end{prop}
\subsection{Equivalence $\mathcal{L}^{k}_{\infty,\omega}(\mathsf{Cnt})$}
The $k$-variable bijective pebbling game between two pure $\sigma'$-structures $A,B$, $\pgameBij{k}{A}{B}$ is played by Spoiler and Duplicator as follows:
\begin{itemize}
\item If cardinality of $A$ is not equal to the cardinality of $B$, then Spoiler wins the game.
\item If cardinality of $A$ is equal to the cardinality of $B$, then for each round $i$, Spoiler places a pebble $p$ on element $a_{p}$ and Duplicator selects a bijection $\psi_{i,p}:A \longrightarrow B$. 
\item Duplicator wins the $i$-th round if for $b_{q} = \psi_{i,p}(a_{q})$ where $a_{q}$ is the element in $A$ with pebble $q$, the map $\gamma:a_{q} \mapsto b_{q}$ is a partial isomorphism form $A$ to $B$. 
\end{itemize}
Duplicator wins the game if Duplicator wins the $i$-th round for all $i \in \omega$. The game $\pgameBij{k}{A}{B}$ characterizes equivalence in the $k$-variable fragment of $\mathcal{L}_{\omega,\infty}(\mathsf{Cnt})$. Namely,
\begin{prop}[{\cite{Hella1996}}]
The following are equivalent:
\begin{itemize}
\item Duplicator has a winning strategy in $\pgameBij{k}{A}{B}$
\item $\equivL{A}{B}{\mathcal{L}^{k}_{\infty,\omega}(\mathsf{Cnt})}$
\end{itemize}
\label{prop:bijToSyntaxP}
\end{prop}
Our goal is to show that isomorphism in the coKleisli category internalizes $\pgameBij{k}{A}{B}$. Suppose $f:\pcomonad{k}{A} \longrightarrow B$ is a $\sigma$-morphism. For all $s \in \pcomonad{k}{A} \cup \{\emptySeq\}$, define the function $\psi_{s,p}:A \longrightarrow B$ by $a \mapsto f(s[(p,a)])$. Using these functions and the following definition we can show that isomorphism in the coKleisli category internalizes $\pgameBij{k}{A}{B}$ and characterizes equivalence in $\mathcal{L}^{k}_{\infty,\omega}$   
\begin{defn}
Given $f:\pcomonad{k}{A} \longrightarrow B$, $f$ is \textit{winning for Duplicator in $\pgameBij{k}{A}{B}$} if $\psi_{s,p}$ bijective for all $s \in \pcomonad{k}{A} \cup \{\emptySeq\}$ and $p \in [k]$.  
\end{defn}
\begin{prop}[{\cite[Theorem 15]{Abramsky2017}}]
$f:\pcomonad{k}{A} \longrightarrow B$ is an isomorphism in $\mathcal{K}(\pcomonad{k}{})$ if and only if $f$ is winning for Duplicator in $\pgameBij{k}{A}{B}$.
\begin{proof}
The proof is essentially the same as in (\ref{prop:bijEF}). 
\end{proof}
\begin{cor}
$A,B$ are isomorphic in $\mathcal{K}(\pcomonad{k}{})$ if and only if $\equivL{A}{B}{\mathcal{L}^{k}_{\infty,\omega}(\mathsf{Cnt})}$.
\begin{proof}
Follows from (\ref{prop:bijToSyntaxP}) and (\ref{prop:bijP})
\end{proof}
\end{cor}
\label{prop:bijP}
\end{prop}
\section{Coalgebras and Tree-Width}
In this chapter, we give another proof of the result \cite[Theorem 23]{Abramsky2017} characterizing tree-width of a structure in terms of coalgebras $\alpha:A \longrightarrow \pcomonad{k}{A}$. The are many equivalent ways to define the notion of tree-width for a $\sigma$-structure $A$. We use the one given \cite{Parra1997} in terms of chordal completions. This proof combines basic facts about chordal graphs and the correspondence in \cite[Theorem 6]{Abramsky2017}. It has a very similar structure to the proof of (\ref{cor:treeDepth}) and exposes the similarity between $\efcomonad{k}{}$ and $\pcomonad{k}{}$.  
\begin{defn}[{\cite{Fulkerson1965}}]
A \textit{chordal graph} $\mathcal{C}$ is a undirected graph such that there exists a linear order $<_{\mathcal{C}}$ on the vertices of $\mathcal{C}$ with all $\leq_{\mathcal{C}}$-predecessors of $v \in \mathcal{C}$ adjacent to $v$ form a clique. The ordering $<_{\mathcal{C}}$ is called a perfect elimination ordering of $\mathcal{C}$.   
\label{defn:chordalGraph}
\end{defn}
\begin{defn}
Given a undirected graph $G$, a chordal completion of $G$ is a chordal graph $\mathcal{C}$ such that $G$ is a subgraph of $\mathcal{C}$ 
\end{defn}
\begin{defn}
Given an undirected graph $G$, the \textit{clique number of $G$} is the maximum number of vertices of a clique in $G$.
\end{defn}
Recall the definition (\ref{defn:gaifmanGraph}) of the Gaifman graph $\mathcal{G}(A)$ of a $\sigma$-structure $A$.
\begin{defn}[{\cite{Parra1997}}]
Given a $\sigma$-structure $A$, the \textit{tree-width} of $A$, denoted $\mathsf{tw}(A)$, is one less than least clique number of a chordal completion $\mathcal{C}$ for $\mathcal{G}(A)$. 
\label{defn:treeWidth}
\end{defn}
\begin{prop}
Let $A$ be finite $\sigma$-structure. There is a bijective correspondence between:
\begin{enumerate}[label=(\arabic*)]
\item coalgebras $\alpha:A \longrightarrow \pcomonad{k}{A}$
\item Chordal completions $\mathcal{C}$ of $\mathcal{G}(A)$ with clique number $k$
\end{enumerate}
\begin{proof}
For the $(1) \Rightarrow (2)$ correspondence, suppose $\alpha:A \longrightarrow \pcomonad{k}{A}$ is a coalgebra. Construct a chordal graph $\mathcal{C}$ with vertices being the elements of $A$. By unpacking (\ref{eq:coalgebraLaw}-left), $\alpha$ gives a tree-order on $A$ where $a \leq a' \Leftrightarrow \alpha(a) \sqsubseteq \alpha(a')$. Choose a root for this tree and let $\leq_{\mathcal{C}}$ be the preorder traversal of the $\leq$-tree order. Let there be an edge $(a,a')$ in $\mathcal{C}$ if $\alpha(a) \sqsubseteq \alpha(a')$ or $\alpha(a') \sqsubseteq \alpha(a)$.   
It is clear by construction of $\mathcal{C}$, that $\leq_{\mathcal{C}}$ is a perfect elimination ordering and $\mathcal{C}$ is a chordal graph. Suppose $(a,a')$ is an edge in $\mathcal{G}(A)$, then by the definition of Gaifman graph there exists an $R \in \sigma$ such that $a$ and $a'$ appear in a tuple of $R^{A}$. By $\alpha$ a $\sigma$-morphism, $\alpha(a)$ and $\alpha(a')$ appear in a tuple in $R^{\pcomonad{k}{A}}$. By the interpretation of $R$ on $\pcomonad{k}{A}$, $\alpha(a) \sqsubseteq \alpha(a')$ or $\alpha(a') \sqsubseteq \alpha(a)$. Without loss of generality, assume that $\alpha(a) \sqsubseteq \alpha(a')$. Hence, there is edge $(a,a')$ in $\mathcal{C}$. Therefore, $\mathcal{G}(A)$ is subgraph $\mathcal{C}$. Moreover, since $\alpha(a)$, $\alpha(a')$ appear in the same tuple of some $R^{\pcomonad{k}{A}}$, the pebble index of $a$ in $\alpha(a')$ must not occur in the suffix of $\alpha(a)$ in $\alpha(a')$. It follows that there at most $k$ elements adjacent to $a'$ in $\mathcal{C}$ that are $\leq_{\mathcal{C}}$-predecessors of $a'$. Hence, $\mathcal{C}$ is a chordal completion of $\mathcal{G}(A)$ with clique number $k$. \\~\\
For the $(2) \Rightarrow (1)$ correspondence, suppose $\mathcal{C}$ is a chordal completion of $\mathcal{G}(A)$ with clique number $k$. By chordal graphs, being perfect graphs, every induced subgraph has clique number equal to its chromatic number (i.e. least number of colours for a valid colouring). In particular, the graph $\mathcal{C}$ itself has a $k$-colouring $i:A \longrightarrow [k]$. By the definition of chordal graph (\ref{defn:chordalGraph}), $\mathcal{C}$ has a perfect elimination ordering $<_{\mathcal{C}}$. Define $\alpha(a) = [(i(a_{1}),a_{1}),\dots,(i(a_{n}),a_{n})]$ where $a_{1},\dots,a_{n}$ are the $\leq_{\mathcal{C}}$-predecessors of $a$ (including $a = a_{n}$ itself) adjacent to $a$ in $\mathcal{C}$. Clearly, by definition the counit law (\ref{eq:coalgebraLaw}-right) is satisfied. Since $\leq_{\mathcal{C}}$ is a linear order, the $\leq_{\mathcal{C}}$-predecessors are unique, so the other coalgebra law (\ref{eq:coalgebraLaw}-left) is satisfied.       
\end{proof}
\label{prop:coalgebraChordal}
\begin{defn}
Given a $\sigma$-structure $A$, define the \textit{$\mathbb{T}$-coalgebra number of $A$}, denoted $\kappa(A)$, as the least $k$ such that there exists a coalgebra $\alpha:A \longrightarrow \pcomonad{k}{A}$. 
\end{defn}
\begin{cor}
$\kappa(A) = \mathsf{tw}(A) + 1$
\begin{proof}
This follows directly from the proposition (\ref{prop:coalgebraChordal}) and definition (\ref{defn:treeWidth}). The proof has the exact same structure as the proof of (\ref{cor:treeDepth}).   
\end{proof}
\label{cor:treeWidth}
\end{cor}
\end{prop}
Hence, as the above corollary (\ref{cor:treeWidth}) shows, the graded family of comonads $\{\pcomonad{k}{}\}_{k \in \omega}$ gives a purely categorical definition of tree-width of a structure. Namely, the tree-width of a structure $A$ is $k-1$ where $k$ is least number such that a coalgebra $\alpha:A \longrightarrow \pcomonad{k}{A}$ exists.
